---
title: "LAGOS Spatial Analysis"
author: "Austin Hopkins"
date: "2/23/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
#install.packages("mapview")
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(RApiSerialize)
library(lubridate)
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
# #Lagos download script
#LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus


load('lake_centers.Rdata')


```



### Convert to spatial data
```{r}
#Look at the column names
names(lake_centers)

#Look at the structure
str(lake_centers)

#View the full dataset
View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
#Iowa
iowa <- states%>%
  filter(name== 'Iowa')%>%
  st_transform(2163)

#Subset Iowa out
Iowa_lakes <- spatial_lakes[iowa,]

#Plotting the first 1000 lakes in Iowa
Iowa_lakes%>%
  arrange(-lake_area_ha)%>%
  slice(1:1000)%>%
  mapview(.,zcol='lake_area_ha')

#Illinois
illinois <- states %>%
  filter(name== 'Illinois')%>%
  st_transform(2163)

#Subset Iowa out
illinois_lakes <- spatial_lakes[illinois,]

#Plotting the first 1000 lakes in Iowa
illinois_lakes%>%
  arrange(-lake_area_ha)%>%
  slice(1:1000)%>%
  mapview(.,zcol='lake_area_ha')

#I need to combine the two states
Iowa_Illinois <- states %>%
  filter(name == 'Iowa'| name =='Illinois') %>%
  st_transform(2163)
mapview(Iowa_Illinois)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

```{r}
#pull out only Ia and Il states

ia_il_lakes <- spatial_lakes[Iowa_Illinois,] %>% 
  group_by(state_zoneid) %>% 
  summarize(ia_il_lakes)
View(ia_il_lakes)
```
in Minnesota there are give or take 29,038 lakes. In Illinois and Iowa together there are a meger 16,466 lakes. Maybe if Iowa and Illinois got their act together they could be more like **Minnesota**.


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
#Lets make a histy!
ggplot(data = ia_il_lakes) +
    geom_histogram(aes(x=lake_area_ha), binwidth = 1500)

ia_il_lakes %>%
  arrange(-lake_area_ha) %>%
  slice(1:1000) %>%
  ggplot() +
    geom_histogram(aes(x=lake_area_ha), binwidth = 1500)


```

Most of the lakes in question are teeny tiny, with an area less than 2500 hectares. There are a handful of more substantial lakes that make the histogram impossible to view effectively. The large lakes were cut out and only the 1000 smallest lakes were kept. 



## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}
ia_il_lakes %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 

Adding a depth measurement variable to this dataset would broaden our data. a small surface area lake can be incredibly deep. Some sort of flow metric as well, tracking water gain and loss from up and to downstream. I would potenntially add a precipitation factor in here as well as the watersheds ability to hold onto water in the soil. 








